---
title: Accidents analysis
author: Alvaro
date: '2024-01-10'
slug: new-post-dudes
categories:
  - R
tags:
  - plot
subtitle: ''
summary: ''
authors: []
lastmod: '2024-01-10T16:07:43-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## The data

We’re going to explore data from the National Electronic Injury Surveillance System (NEISS), collected by the Consumer Product Safety Commission.

It’s an interesting dataset to explore because every one is already familiar with the domain, and each observation is accompanied by a short narrative that explains how the accident occurred.

```{r load data, message=FALSE}
library(tidyverse)

```

```{r injuries, message=FALSE}
injuries <- read_tsv("C:/Users/admin/Desktop/R/R-todo/ER_injuries/neiss/injuries.tsv")
spec(injuries)


```

```{r products, message=FALSE}
products <- read_tsv("C:/Users/admin/Desktop/R/R-todo/ER_injuries/neiss/products.tsv")
spec(products)

```

```{r pop, message=FALSE}
population <- read_tsv("C:/Users/admin/Desktop/R/R-todo/ER_injuries/neiss/population.tsv")
spec(population)
```

## Exploration 

We’ll start by looking at a product with an interesting story: 649, “toilets”. First, we’ll pull out the injuries associated with this product:

```{r}
selected <- injuries %>% filter(prod_code == 649)
nrow(selected)

```

Next, we’ll perform some basic summaries looking at the location, body part, and diagnosis of toilet-related injuries.

```{r}
selected %>% count(location, wt = weight, sort = TRUE)
```

```{r}
selected %>% count(body_part, wt = weight, sort = TRUE)
```

```{r}
selected %>% count(diag, wt = weight, sort = TRUE)
```

As you might expect, injuries involving toilets most often occur at home. The most common body parts involved possibly suggest that these are falls (since the head and face are not usually involved in routine toilet usage), and the diagnoses seem rather varied.

We can also explore the pattern across age and sex. We have enough data here that a table is not that useful, and so I make a plot, Fig..., that makes the patterns more obvious.

```{r}
summary <- selected %>% 
  count(age, sex, wt = weight)
summary

summary %>% 
  ggplot(aes(age, n, colour = sex)) + 
  geom_line() + 
  labs(y = "Estimated number of injuries")
```

We see a spike for young boys peaking at age 3, and then an increase (particularly for women) starting around middle age, and a gradual decline after age 80.

One problem with interpreting this pattern is that we know that there are fewer older people than younger people, so the population available to be injured is smaller. We can control for this by comparing the number of people injured with the total population and calculating an injury rate. Here I use a rate per 10,000.

```{r}
summary <- selected %>% 
  count(age, sex, wt = weight) %>% 
  left_join(population, by = c("age", "sex")) %>% 
  mutate(rate = n / population * 1e4)

summary


```

```{r}
summary %>% 
  ggplot(aes(age, rate, colour = sex)) + 
  geom_line(na.rm = TRUE) + 
  labs(y = "Injuries per 10,000 people")
```
